Statically Typed String Sanitation Inside a Python

### 2014-08-18 sanitation-psp14/sanitation-psp14-short.tex
Monday meeting changes #1
 * Fixed semantics (audit the check rules)
 * Fixed names of thms/lems/defs and changed lems to props in sec 2

\inferrule[S-E-App]
{ e_1 \sreduces e_1' \\ e_2 \sreduces e_2' }
{ e_1(e_2) \sreduces e_1'(e_2')}

@@

\inferrule[S-E-Check-Ok]
{e \sreduces \strin{s} \\ s \in \lang{r}}
{
  \rcheck{r}{e}{x.e_1}{e_2} \sreduces [\strin{s} / x]e_1
}

@@

\inferrule[P-E-App]
{ \iota_1 \sreduces \iota_1' \\  \iota_2 \sreduces \iota_2' }
{ \iota_1(\iota_2) \sreduces \iota_1'(\iota_2')}

@@

\inferrule[P-E-Check-OK]
{\iota \treduces \tstr{s} \\ s \in \lang{r}}
{\tcheck{\rx{r}}{\iota}{x.\iota_1}{\iota_2} \treduces [\tstr{s}/x]\iota_1}

@@

\begin{lem}{Properties of Regular Languages and Expressions.} \label{thm:regexprops}
The following are properties of regular expressions which are necessary for our proofs:
If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $s_1s_2 \in \lang{r_1r_2}$.
For all strings $s$ and regular expressions $r$, either $s \in \lang{r}$ or $s \not \in \lang{r}$.
Regular languages are closed under difference, right quotient, reversal, and string homomorphism.
\end{lem}

@@

\begin{lem}{Closure.} \label{thm:total}
  If $\lang{r}, \lang{r_1}$ and $\lang{r_2}$ are regular expressions, then $\lang{\lreplace{r}{r_1}{r_2}}$ is also a regular language.
\end{lem}
\begin{proof}
The theorem follows from closure under difference, right quotient, reversal and string homomorphism.
\end{proof}

\begin{lem}{Substitution Correspondence.} \label{thm:substcorrespondence}
  If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{s_1}{s_2}}$.
\end{lem}
\begin{proof}
The theorem follows from the definitions of subst and lreplace; note that language substitutions over-approximate string substitutions.
\end{proof}

@@

\begin{thm}{Type Safety for the String Fragment of $\lambdas$.} \label{thm:scorrect}
  Let $S$ be a term in the source language. If $\sctx \vdash e : \stringin{r}$ 
  then $e \sreduces \rstr{s}$ and $\sctx \vdash \rstr{s} : \stringin{r}$.
\end{thm}
\begin{proof}
By induction on the typing relation.
The S-T-Concat case requires Lemma \ref{thm:regexprops} 
and the S-T-Replace case appeals to Lemma \ref{thm:substcorrespondence}.
\end{proof}

@@

\begin{thm}{Correctness of Input Sanitation for $\lambdas$.}\label{thm:scorrect}
  If  $\sctx \vdash e : \stringin{r}$ and $e \sreduces \rstr{s}$ then $s \in \lang{r}$.
\end{thm}
\begin{proof}
  Follows directly from type safety, canonical forms for $\lambdas$.
\end{proof}

@@

\begin{thm}{Translation Correctness.} \label{thm:tcorrect}
If $\sctx \vdash e : \stringin{r}$ then there exists a $\iota$ such that $\trden{e} = \iota$, 
$\iota \treduces \tstr{s}$, and $e \sreduces \strin{s}$.
\end{thm}

@@

\begin{thm}{Correctness of Input Sanitation for Translated Terms.}\label{thm:main}

>>>>

\inferrule[S-E-App]
{ e_1 \sreduces \lambda x . e_3 \\ e_2 \sreduces e_2' \\ [ e_2' / x ] e_3 \sreduces v }
{ e_1(e_2) \sreduces v }

@@

\inferrule[S-E-Check-Ok]
{e \sreduces \strin{s} \\ s \in \lang{r} \\ [\strin{s} / x]e_1 \sreduces v}
{
  \rcheck{r}{e}{x.e_1}{e_2} \sreduces v
}

@@

\inferrule[P-E-App]
{ \iota_1 \sreduces \lambda x . \iota_3 \\  \iota_2 \sreduces \iota_2' \\ [\iota_3 / x] \iota_2' \sreduces v}
{ \iota_1(\iota_2) \sreduces v}

@@

\inferrule[P-E-Check-OK]
{\iota \treduces \tstr{s} \\ s \in \lang{r} \\ [\tstr{s} / x] \iota_1 \sreduces \iota_3}
{\tcheck{\rx{r}}{\iota}{x.\iota_1}{\iota_2} \treduces \iota_3}

@@

\begin{lem}[Properties of Regular Languages and Expressions.] \label{thm:regexprops}
The following are properties of regular expressions which are necessary for our proofs:
If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $s_1s_2 \in \lang{r_1r_2}$.
For all strings $s$ and regular expressions $r$, either $s \in \lang{r}$ or $s \not \in \lang{r}$.
Regular languages are closed under difference, right quotient, reversal, and string homomorphism.
\end{lem}

@@

\begin{prop}[Closure.] \label{thm:total}
  If $\lang{r}, \lang{r_1}$ and $\lang{r_2}$ are regular expressions, then $\lang{\lreplace{r}{r_1}{r_2}}$ is also a regular language.
\end{prop}
\begin{proof}
The theorem follows from closure under difference, right quotient, reversal and string homomorphism.
\end{proof}

\begin{prop}[Substitution Correspondence.] \label{thm:substcorrespondence}
  If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{s_1}{s_2}}$.
\end{prop}
\begin{proof}
The theorem follows from the definitions of subst and lreplace; note that language substitutions over-approximate string substitutions.
\end{proof}

@@

\begin{thm}[Type Safety for the String Fragment of $\lambdas$.] \label{thm:scorrect}
  If $\sctx \vdash e : \stringin{r}$ 
  then $e \sreduces \rstr{s}$ and $\sctx \vdash \rstr{s} : \stringin{r}$.
\end{thm}
\begin{proof}
By induction on the typing relation.
The S-T-Concat case requires Lemma \ref{thm:regexprops} 
and the S-T-Replace case appeals to Lemma \ref{thm:substcorrespondence}.
\end{proof}

@@

\begin{thm}[Correctness of Input Sanitation for $\lambdas$.]\label{thm:scorrect}
  If  $\sctx \vdash e : \stringin{r}$ and $e \sreduces \rstr{s}$ then $s \in \lang{r}$.
\end{thm}
\begin{proof}
  Follows directly from type safety, canonical forms for $\lambdas$.
\end{proof}

@@

\begin{thm}[Translation Correctness.] \label{thm:tcorrect}
If $\sctx \vdash e : \stringin{r}$ then there exists a $\iota$ such that $\trden{e} = \iota$, 
$\iota \treduces \tstr{s}$, and $e \sreduces \strin{s}$.
\end{thm}

@@

\begin{thm}[Correctness of Input Sanitation for Translated Terms.]\label{thm:main}
